Tight Generalization Bounds for Hard Parameter Sharing in High-dimensional Linear Regression

### 2020-10-07 arxiv/experiments.tex
finish section 5

\paragraph{Sample ratio.}
First, we validate that as we increase the sample ratio while keeping task two's sample size fixed, task two's prediction accuracy does not always increase.
Figure \ref{fig_ab_data} shows the result for three example task pairs. We fix task one's sample size as $1000$ and increase task two's number of samples from $100$ to $3000$.

Next, we propose a natural progressive training schedule for finding the optimal sample ratio.
Our training schedule is motivated by Corollary \ref{cor_MTL_loss}, which shows that as we increase the sample ratio, the prediction loss always decreases first and then increases.
%In particular, Figure \ref{fig_size} (and our analysis) shows that $L(\hat{\beta}_t^{\MTL})$ behaves as a quadratic function over $\rho_1$.
%More generally, depending on how large $\Psi(\beta_1, \beta_2)$ is, $L(\hat{\beta}_t^{\MTL})$ may also be monotonically increasing or decreasing.
Based on this observation, we propose a progressive training schedule to improve the compuational efficiency of hard parameter sharing.
\begin{itemize}
	\item We divide the source task data into $S$ batches.
	For $S$ rounds, we incrementally add the source task data by adding one batch at a time.
	\item After training $T$ epochs, if the validation accuracy becomes worse than the previous round's result, we terminate.
	Algorithm \ref{alg_inc_train} in Appendix \ref{app_experiments} describes the procedure in detail.
\end{itemize}

@@

%We fill in the details of the experimental procedure used for the results in Figure \ref{fig_ablation}.
%\squishlist
%	\item Task similarity: We select a similar and a dissimilar source task compared to the target task using domain knowledge.
%First pair: the customer review dataset (CR) , which predicts whether a review is positive or negative, is more similar to SST (sentiment treebank) than MPQA (question type).
%Second pair: SST is more similar to MR since they both concern about positive or negative opinions expressed the text.
%TREC is less similar to MR because the task is about question types.
%Third pair: MPQA (opinion polarity) is more similar to TREC (question type)
Our results show that progressive training requires less computation compared to a baseline round-robin training schedule, without sacrificing accuracy.
First, we show that our proposed progressive training schedule (Algorithm \ref{alg_inc_train}) can help mitigate negative transfer for predicting a particular target task.
Over all $15$ pairs from the sentiment analysis tasks, we find that Algorithm \ref{alg_inc_train} requires only $45\%$ of the computational cost to achieve similar performance on the target task, compared to the MTL baseline.
Our insight is that since adding more samples from the source task does not always help, we can improve efficiency by adding source samples \textit{incrementally} during training.
%\textbf{Improving transfer learning training efficiency.}
%We show that Algorithm \ref{alg_inc_train} also applies to transfer learning settings.
%Compared to fine-tuning the source model on the target task, we show that our proposed method reduces the computional cost by \alert{$xx\%$}, without sacrificing accuracy.
Second, we show the progressive training schedule applies to multiple tasks as well.
In Table \ref{tab:taskonomy}, we find that over all six sentiment analysis tasks, incremental training requires less than $35\%$ of the computational cost compared to baseline MTL training, while achieving the same accuracy averaged over all six tasks.
As a further validation, excluding TREC, we observe similar comparative results.
%The data efficiency ratio of using MLP is $100\%$ because the average performance of MTL is worse than the average of STL.
%We further show that applying incremental training helps reduce the data efficiency ratio to \alert{$xx\%$}.
%If TREC is not included, we see that only $25\%$ of the labeled data is needed.

\begin{figure}%[!t]
	\begin{subfigure}[t]{0.5\textwidth}
		\centering
		\vspace{0pt}
		\includegraphics[width=0.6\textwidth]{figures/ratio_norm_3_pairs_lstm.pdf}
		\caption{HPS vs. STL}
		\label{fig_ab_data}
	\end{subfigure}\hfill
	\begin{subfigure}[t]{0.5\textwidth}
		\centering
		\vspace{0pt}
		\includegraphics[width=0.6\textwidth]{figures/ratio_alignment_norm_diff_all.pdf}
		\caption{HPS vs. covariance alignment}
		\label{fig_ab_cov}
	\end{subfigure}
	\caption{A}
%	\begin{minipage}[t]{0.4\textwidth}
%		\centering
%		\vspace{0pt}
%		\begin{tabular}{c c c}
%		\toprule
		% \multirow{2}{*}{{\bf Models}} & \multicolumn{2}{c}{\begin{minipage}{1.2in}\begin{center}
		% Sentiment\\ analysis\end{center}\end{minipage}} \\
%		\multirow{2}{*}{{\bf Models}} & \multicolumn{2}{c}{\bf Sentiment analysis} \\
		% \cmidrule(lr){2-3}
%		& all tasks & w/o TREC \\
%		\midrule
%		{\bf MLP}  & 31\% & 29\% \\
%		{\bf LSTM} & 35\% & 34\% \\
%		{\bf CNN}  & 30\% & 28\% \\
%		\bottomrule
%		\end{tabular}
%		\captionof{table}{Finding the best sample ratio via a progressive training schedule.}
%		\label{tab:taskonomy}
%	\end{minipage}
\end{figure}

%	\begin{subfigure}[b]{0.33\textwidth}
%		\centering
%		\includegraphics[width=0.975\textwidth]{figures/task_sim_norm_lstm.pdf}
%		\caption{Task similarity}
%		\label{fig_ab_sim}
%	\end{subfigure}%
%	\caption{Validating the three results of Section \ref{sec_insight} on sentiment analysis tasks. (a) Adding a semantically similar source task in MTL performs better than adding a dissimilar task.
%	(b) As source/target sample ratio increases, we observe a transition from positive to negative transfer.
%	(c) As source/target sample ratio increases, aligning task covariances \cite{WZR20} improves
%more over the baseline.
%	Note: (S) denotes the source task and (T) denotes the target task.}

@@

\paragraph{Covariate shift.}
Our observation highlights the need to correct covariate shift when $n_1 / n_2$ is large.
Hence for such settings, we expect procedures that aim at correcting covariate shift to provide more significant gains.
We consider a covariance alignment procedure proposed in \cite{WZR20}, which is designed for the purpose of correcting covariate shift.
The idea is to add an alignment module between the input and the shared module $B$.
This new module is then trained together with $B$ and the output layers.
We validate our insight on this procedure in the experiments.
We implement the covariance alignment procedure following \cite{WZR20}.
We fix the sample size of the target task as $1000$.

In Figure \ref{fig_ab_cov}, we measure the performance gains from performing the alignment procedure proposed in \cite{WZR20} minus baseline MTL.
We average the results over all 15 task pairs.
The result shows that as the source samples increases, the alignment procedure shows a bigger improvement over MTL.

>>>>

\paragraph{Sample ratio.}
In Figure \ref{fig_ab_data}, we observe that for multiple task pairs, increasing task one's sample size improves task two's prediction accuracy initially, but hurts eventually.
On the $y$-axis, we plot task two's prediction accuracy using HPS, subtracted by its STL prediction accuracy.
%validate that as we increase the sample ratio while keeping task two's sample size fixed, task two's prediction accuracy does not always increase.
We fix task one's sample size as $1000$ and increase task two's number of samples from $100$ to $3000$.

Next, we propose a natural progressive training schedule for finding the optimal sample ratio.
Our training schedule is motivated by Corollary \ref{cor_MTL_loss}, which shows that as we increase the sample ratio, the prediction loss always decreases first and then increases.
%In particular, Figure \ref{fig_size} (and our analysis) shows that $L(\hat{\beta}_t^{\MTL})$ behaves as a quadratic function over $\rho_1$.
%More generally, depending on how large $\Psi(\beta_1, \beta_2)$ is, $L(\hat{\beta}_t^{\MTL})$ may also be monotonically increasing or decreasing.
Based on this observation, we propose a progressive training schedule to improve the compuational efficiency of hard parameter sharing.
\begin{itemize}
	\item We divide the source task data into $S$ batches.
	For $S$ rounds, we incrementally add the source task data by adding one batch at a time.
	\item After training $T$ epochs, if the validation accuracy becomes worse than the previous round's result, we terminate.
%	Algorithm \ref{alg_inc_train} in Appendix \ref{app_experiments} describes the procedure in detail.
\end{itemize}

@@

%We fill in the details of the experimental procedure used for the results in Figure \ref{fig_ablation}.
%\squishlist
%	\item Task similarity: We select a similar and a dissimilar source task compared to the target task using domain knowledge.
%First pair: the customer review dataset (CR) , which predicts whether a review is positive or negative, is more similar to SST (sentiment treebank) than MPQA (question type).
%Second pair: SST is more similar to MR since they both concern about positive or negative opinions expressed the text.
%TREC is less similar to MR because the task is about question types.
%Third pair: MPQA (opinion polarity) is more similar to TREC (question type)
Our results show that progressive training requires less computation compared to a baseline round-robin training schedule, without sacrificing accuracy.
First, we show that our proposed progressive training schedule (can help mitigate negative transfer for predicting a particular target task.
Over all $15$ pairs from the sentiment analysis tasks, we find that progressive training requires only $45\%$ of the computational cost to achieve the same test accuracy on the target task, compared to HPS trained using round-robin.
Our insight is that since adding more samples from the source task does not always help, we can improve efficiency by adding source samples \textit{progressively} during training.
%\textbf{Improving transfer learning training efficiency.}
%We show that Algorithm \ref{alg_inc_train} also applies to transfer learning settings.
%Compared to fine-tuning the source model on the target task, we show that our proposed method reduces the computional cost by \alert{$xx\%$}, without sacrificing accuracy.
Second, we show the progressive training schedule applies to multiple tasks as well.
Over all six sentiment analysis tasks, incremental training requires less than $35\%$ of the computational cost compared to baseline MTL training, while achieving the same accuracy averaged over all six tasks.
%As a further validation, excluding TREC, we observe similar comparative results.
%The data efficiency ratio of using MLP is $100\%$ because the average performance of MTL is worse than the average of STL.
%We further show that applying incremental training helps reduce the data efficiency ratio to \alert{$xx\%$}.
%If TREC is not included, we see that only $25\%$ of the labeled data is needed.

\begin{figure}%[!t]
	\begin{subfigure}[t]{0.5\textwidth}
		\centering
		\vspace{0pt}
		\includegraphics[width=0.6\textwidth]{figures/ratio_norm_3_pairs_lstm.pdf}
		\caption{HPS vs. STL}
		\label{fig_ab_data}
	\end{subfigure}\hfill
	\begin{subfigure}[t]{0.5\textwidth}
		\centering
		\vspace{0pt}
		\includegraphics[width=0.6\textwidth]{figures/ratio_alignment_norm_diff_all.pdf}
		\caption{HPS vs. covariance alignment}
		\label{fig_ab_cov}
	\end{subfigure}
	\caption{Comparing hard parameter sharing to single-task learning (STL) and a covariance alignment approach proposed by \citet{WZR20}.
	In Figure \ref{fig_ab_data}, we observe that for multiple example task pairs, increasing task one's sample size improves task two's prediction accuracy initially, but hurts eventually -- a phenomenon similar  to Figure \ref{fig_size}.
	In Figure \ref{fig_ab_cov}, we observe that as task one's sample size increases, covariance alignment improves more over HPS. }
	\label{fig_text}
%	\begin{minipage}[t]{0.4\textwidth}
%		\centering
%		\vspace{0pt}
%		\begin{tabular}{c c c}
%		\toprule
		% \multirow{2}{*}{{\bf Models}} & \multicolumn{2}{c}{\begin{minipage}{1.2in}\begin{center}
		% Sentiment\\ analysis\end{center}\end{minipage}} \\
%		\multirow{2}{*}{{\bf Models}} & \multicolumn{2}{c}{\bf Sentiment analysis} \\
		% \cmidrule(lr){2-3}
%		& all tasks & w/o TREC \\
%		\midrule
%		{\bf MLP}  & 31\% & 29\% \\
%		{\bf LSTM} & 35\% & 34\% \\
%		{\bf CNN}  & 30\% & 28\% \\
%		\bottomrule
%		\end{tabular}
%		\captionof{table}{Finding the best sample ratio via a progressive training schedule.}
%		\label{tab:taskonomy}
%	\end{minipage}
\end{figure}

%	\begin{subfigure}[b]{0.33\textwidth}
%		\centering
%		\includegraphics[width=0.975\textwidth]{figures/task_sim_norm_lstm.pdf}
%		\caption{Task similarity}
%		\label{fig_ab_sim}
%	\end{subfigure}%
%	\caption{Validating the three results of Section \ref{sec_insight} on sentiment analysis tasks. (a) Adding a semantically similar source task in MTL performs better than adding a dissimilar task.
%	(b) As source/target sample ratio increases, we observe a transition from positive to negative transfer.
%	(c)
%	Note: (S) denotes the source task and (T) denotes the target task.}

@@

\paragraph{Covariate shift.}
Recall from Example \ref{ex_covshift} that when the sample ratio increases, covariate shift worsens.
Hence when the sample ratio increases, we expect procedures that aim at correcting covariate shift to provide more significant gains.
To this end, we study a covariance alignment procedure proposed in \citet{WZR20}, which is designed for correcting covariate shift.
The idea is to add an alignment module between the input and the shared module $B$.
This new module is then trained together with $B$ and the output layers -- more details on the procedure and the implementation can be found in that paper.
%We validate our insight on this procedure in the experiments.
%We implement the covariance alignment procedure following \cite{WZR20}.

In Figure \ref{fig_ab_cov}, we measure the performance gains from performing covariance alignment vs. HPS.
We average the results over all 15 task pairs.
The result shows that as the source samples increases, covariance alignment improves more over HPS.
We fix task two's sample size at $1,000$ and increase task one's sample size from $1,000$ to $3,000$.
